---
title: "Battery Project"
author: "Bhavin Bohre"
date: "2026-03-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
library(lubridate)
```

```{r}
battery = read.csv("battery_dataset_20260327.csv")
```

```{r}
colSums(is.na(battery))
```

```{r}
battery[!complete.cases(battery), ]
```

```{r}
battery = battery[!(is.na(battery$Quantity)), ]
```

```{r}
account_a = subset(battery, AccountNumber == "account_a")

account_b = subset(battery, AccountNumber == "account_b")
```

```{r}
ggplot(account_a, aes(y = Quantity)) +
geom_boxplot(fill = "royalblue", alpha = 0.6) +
labs(title = "Distribution of Quantity", y = "Quantity") +
theme_minimal()
```


```{r}
ggplot(account_a, aes(x = Quantity)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Quantity", x = "Quantity", y = "Density") +
theme_minimal()
```

```{r}
# Calculate interquartile range for quantity
Q1 = quantile(account_a$Quantity, 0.25, na.rm = TRUE)
Q3 = quantile(account_a$Quantity, 0.75, na.rm = TRUE)
IQR_value = IQR(account_a$Quantity, na.rm = TRUE)

# Set the bounds for outliers
lower_bound = Q1 - 1.5 * IQR_value
upper_bound = Q3 + 1.5 * IQR_value

# Remove rows that have outlier values in quantity column
account_a = subset(account_a, Quantity >= lower_bound & Quantity <= upper_bound)
```

```{r}
ggplot(account_a, aes(y = Quantity)) +
geom_boxplot(fill = "royalblue", alpha = 0.6) +
labs(title = "Distribution of Quantity", y = "Quantity") +
theme_minimal()
```

```{r}
ggplot(account_a, aes(x = Quantity)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Quantity", x = "Quantity", y = "Density") +
theme_minimal()
```
```{r}
account_a$log_quantity = log(account_a$Quantity)
```

```{r}
ggplot(account_a, aes(x = log_quantity)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Log Quantity", x = "Log Quantity", y = "Density") +
theme_minimal()
```

```{r warning = FALSE}
account_a$Datetime = mdy_hm(account_a$Datetime)

account_a$Season = factor(
  case_when(
    month(account_a$Datetime) %in% c(3,4,5) ~ "Spring",
    month(account_a$Datetime) %in% c(6,7,8) ~ "Summer",
    month(account_a$Datetime) %in% c(9,10,11) ~ "Fall",
    TRUE ~ "Winter"
  ),
  levels = c("Spring", "Summer", "Fall", "Winter")
)

ggplot(account_a, aes(x = Datetime, y = Quantity, color = Season)) +
geom_line() + scale_x_datetime(date_breaks = "2 month", date_labels = "%m/%y"
) + labs(title = "Quantity Energy Used vs Season", x = "Date",
y = "Quantity") + theme_minimal()
```

```{r}
# Get the time from the DateTime variable
account_a$Time = format(account_a$Datetime, "%H:%M")

# Create TimeOfDay variable to separate timing of each day
account_a$TimeOfDay = with(account_a, ifelse(
  Time > "05:00" & Time <= "12:00", "Morning",
  ifelse(Time > "12:00" & Time <= "17:00", "Afternoon",
  ifelse(Time > "17:00" & Time <= "20:00", "Evening",
  "Night"))))

# Look at average energy used for time of day
quant_avg = account_a %>% group_by(TimeOfDay) %>%
summarise(avg_quantity = mean(Quantity, na.rm = TRUE))

quant_avg <- quant_avg %>% drop_na(TimeOfDay)

ggplot(quant_avg, aes(x = TimeOfDay, y = avg_quantity, fill = TimeOfDay)) +
geom_col() + labs(title = "Average Quantity by Time of Day",
x = "Time of Day", y = "Average Quantity") + theme_minimal()
```

```{r}
# Plot energy used by time of day
ggplot(account_a, aes(x = Datetime, y = Quantity)) +
geom_line(color = "steelblue") + scale_x_datetime(date_breaks = "2 month",
date_labels = "%m/%y") +
facet_wrap(~ TimeOfDay, ncol = 1, scales = "free_x") +
labs(title = "Quantity vs Time of Day", x = "Date", y = "Quantity"
) + theme_minimal()
```

```{r}
# Convert price from megawatt hours to kilowatt hours
account_a$Settlement.Point.Price.KWH = account_a$SettlementPointPrice / 1000
```

```{r}
ggplot(account_a, aes(y = Settlement.Point.Price.KWH)) +
geom_boxplot(fill = "royalblue", alpha = 0.6) +
labs(title = "Distribution of Price Point", y = "Price Point") +
theme_minimal()
```

```{r}
ggplot(account_a, aes(x = Settlement.Point.Price.KWH)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Price Point", x = "Price Point", y = "Density") +
theme_minimal()
```

```{r}
# Calculate interquartile range for quantity
Q1 = quantile(account_a$Settlement.Point.Price.KWH, 0.25, na.rm = TRUE)
Q3 = quantile(account_a$Settlement.Point.Price.KWH, 0.75, na.rm = TRUE)
IQR_value = IQR(account_a$Settlement.Point.Price.KWH, na.rm = TRUE)

# Set the bounds for outliers
lower_bound = Q1 - 1.5 * IQR_value
upper_bound = Q3 + 1.5 * IQR_value

# Remove rows that have outlier values in quantity column
account_a = subset(account_a, Settlement.Point.Price.KWH >= lower_bound 
& Settlement.Point.Price.KWH <= upper_bound)
```

```{r}
ggplot(account_a, aes(y = Settlement.Point.Price.KWH)) +
geom_boxplot(fill = "royalblue", alpha = 0.6) +
labs(title = "Distribution of Price Point", y = "Price Point") +
theme_minimal()
```

```{r}
ggplot(account_a, aes(x = Settlement.Point.Price.KWH)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Price Point", x = "Price Point", y = "Density") +
theme_minimal()
```

```{r warning = FALSE}
account_a$log_price_point_KWH = log(account_a$Settlement.Point.Price.KWH)
ggplot(account_a, aes(x = log_price_point_KWH)) +
geom_density(fill = "royalblue", alpha = 0.4) +
labs(title = "Density Plot of Log Price", x = "Log Price", y = "Density") +
theme_minimal()
```

```{r}
write.csv(account_a, "account_a.csv", row.names = FALSE)
```


## Optimization Variable:

$x_A$ = Number of batteries needed for account A

$\text{charge}_{A,t,d}$ = charge stored in the batteries for account A at hour $t$ on day $d$ where $t = 0,...,23$ and $d = 1,...,14$ (Starting With 2 Week Period First) 

$y_{b,A,t,d}$ = Whether battery $b$ for account A is charging at time $t$ on day $d$ or not

$e_{A,b}$ = Whether battery $b$ exists for account A

$g_{A,t,d}$ = Quantity energy discharged from the battery at hour $t$ on day $d$

$\text{charge}_{A,d,t}$ = Charge in the battery

$\text{discharge}_{A,d,t}$ = Energy being discharged in the battery

## Constants:

$B^{max} = 30$ = Number of batteries (Set to realistic high value for 2 week period Gurobi will not use all if not needed)

$C_{battery}$ = Amount of money the battery costs (11,959 Tesla Powerwall 3)

$\text{Quantity}_{A,t,d}$ = Quantity of energy used for account A at hour $t$ on day $d$

$\text{Price}_{t,d}$ = Price of energy at hour $t$ on day $d$

$R_{charge}$ = Charge rate of the Tesla Powerwall $3$ in kWh

$R_{discharge}$ = Discharge rate of the Tesla Powerwall $3$ in kWh

## Constraints:

$10.8 * x_A <= \text{CS}_{A,d,t} <= 13.5 * x_A$: Charge state for battery system through 2 weeks must must be at a minimum $80\%$ 

$y_{b,A,t,d} <= e_{A,b}$: Battery can only charge if it exists

$\sum_{b=1}^{B^{\max}} e_{A,b} = x_A$: Number of active batteries is the exact same as the ones needs for account $i$

$g_{A,t,d} = \text{Quantity}_{A,t,d}$: Energy demand must be met by battery usage

$\text{charge}_{A,d,t} <= R_{charge} * x_A$: Charge rate limit

$\text{discharge}_{A,d,t} <= R_{discharge} * x_A$: Discharge rate limit

$13.5 * x_A + \text{charge}_{A,1,0} - \text{discharge}_{A,1,0} = \text{CS}_{A,1,0}$: Initial energy within the first hour of usage

$\text{CS}_{A,d',t'} * x_A + \text{charge}_{A,d,t} - \text{discharge}_{A,d,t} = \text{CS}_{A,d,t}$: Energy in the battery at the current time of the day after the first hour

## Notes:

$d',t'$ = Indicates previous day and previous hour in time

## Objective Function:

$min(\sum_{d = 1}^{14} \sum_{t = 0}^{23} (\text{charge}_{A,d,t} \cdot \text{Price}_{t,d}) + x_A * C_{battery})$
