---
title: "Battery Project"
author: "woessner.21"
date: "2026-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df = read.csv("C:/Users/Caleb Woessner/Downloads/merged_lz_houston_data.csv")
```

```{r}
head(df)
```


```{r, warning=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)


```

```{r}
colnames(df)[colnames(df) == 'Delivery.Date_x'] <- 'Date'

colnames(df)[colnames(df) == 'Settlement.Point.Price'] <- 'Price'

colnames(df)[colnames(df) == 'ACCOUNT_NUMBER'] <- 'Account'

colnames(df)[colnames(df) == 'QUANTITY'] <- 'Quantity'

```

```{r}
head(df)
```
```{r}
any(is.na(df))
sum(is.na(df))
```

```{r}
df$Price = df$Price *1000
```

```{r}
head(df)
```

```{r}
##Split each into their own datasets

df_a <- df %>% filter(Account == "account_a")
df_b <- df %>% filter(Account == "account_b")
```

```{r}
print(mean(df_a$Quantity))
print(mean(df_b$Quantity))
```
```{r}
print(max(df_a$Quantity))
print(max(df_b$Quantity))
```
```{r}
print(mean(df_a$Quantity)/13.5)
print(mean(df_b$Quantity)/13.5)
print(max(df_a$Quantity)/13.5)
print(max(df_b$Quantity)/13.5)
```

We are all splitting up doing our own model on the same Tesla Powerall 3 with 13.5 KWH, Address NA's, change price to KWH not MWH, Do my own EDA. Should try to do this all by Monday. Create Presentation of Everything I have done

Could do: Might be smart to get a table of different batteries other than PowerWall 3 just to have down. Could also look into weather data

Look into Home Energy Profile so that I can market it to residents: #Households is a later issue


Calulate how many max batteries could both accounts be using, look also at average(Caleb)

```{r}
df_a <- df_a %>%
  mutate(
    DateTime = ymd_hms(DateTime),
    Date = as.Date(Date)
  )

df_b <- df_b %>%
  mutate(
    DateTime = ymd_hms(DateTime),
    Date = as.Date(Date)
  )


```


## EDA

```{r}
eda_summary <- function(df) {
  df %>%
    summarise(
      avg_price     = mean(Price, na.rm = TRUE),
      min_price     = min(Price, na.rm = TRUE),
      max_price     = max(Price, na.rm = TRUE),
      avg_quantity  = mean(Quantity, na.rm = TRUE),
      min_quantity  = min(Quantity, na.rm = TRUE),
      max_quantity  = max(Quantity, na.rm = TRUE),
      n_days        = n_distinct(Date)
    )
}

eda_summary(df_a)
eda_summary(df_b)
```

### Account A
```{r}
ggplot(df_a, aes(DateTime, Price)) +
  geom_line() +
  labs(title = "Price over time – Account A", y = "Price ($/kWh)")
```

```{r}
df_a %>%
  group_by(Hour) %>%
  summarise(avg_qty = mean(Quantity, na.rm = TRUE)) %>%
  ggplot(aes(Hour, avg_qty)) +
  geom_line() +
  labs(title = "Average Quantity by Hour – Account A", y = "Quantity")
```

```{r}
df_a %>%
  group_by(Hour) %>%
  summarise(avg_price = mean(Price)) %>%
  ggplot(aes(Hour, avg_price)) +
  geom_line() +
  labs(title = "Average Price by Hour – Account A")
```


```{r}
df_a %>%
  group_by(Month) %>%
  summarise(
    avg_qty   = mean(Quantity, na.rm = TRUE),
    avg_price = mean(Price, na.rm = TRUE)
  ) %>%
  ggplot(aes(Month, avg_qty)) +
  geom_col() +
  labs(title = "Average Monthly Quantity – Account A", y = "Quantity")
```



### Account B

```{r}
ggplot(df_b, aes(DateTime, Price)) +
  geom_line() +
  labs(title = "Price over time – Account B", y = "Price ($/kWh)")
```

```{r}
df_b %>%
  group_by(Hour) %>%
  summarise(avg_qty = mean(Quantity, na.rm = TRUE)) %>%
  ggplot(aes(Hour, avg_qty)) +
  geom_line() +
  labs(title = "Average Quantity by Hour – Account B", y = "Quantity")
```

```{r}
df_b %>%
  group_by(Month) %>%
  summarise(
    avg_qty   = mean(Quantity, na.rm = TRUE),
    avg_price = mean(Price, na.rm = TRUE)
  ) %>%
  ggplot(aes(Month, avg_qty)) +
  geom_col() +
  labs(title = "Average Monthly Quantity – Account B", y = "Quantity")
```

## Ideal Number of Batteries

Doing this based on the Powerall 3 Tesla battery with a capacity of 13.5 kWh

```{r}
battery_cap <- 13.5

batt_stats <- function(df) {
  df <- df %>%
    mutate(batt_equiv = Quantity / battery_cap)
  
  max_batt <- ceiling(max(df$batt_equiv, na.rm = TRUE))
  
  daily_energy <- df %>%
    group_by(Date) %>%
    summarise(total_kwh = sum(Quantity, na.rm = TRUE))
  
  ideal_batt <- mean(daily_energy$total_kwh) / battery_cap
  
  list(
    max_batt   = max_batt,
    ideal_batt = ideal_batt
  )
}

batt_stats(df_a)
batt_stats(df_b)
```
SO the Ideeal battery is the typical number of batteries used. Calculated by finding the average daily energy and deviding it by 13.5
The max is found by what it is during the worst interval of time. So ideal battery is about 1 for a and about 18 for B

This is still Naive but want to get a general idea.


## Simple Naive approach Optimization Model For Model A(account A)

Goal: Charge during cheap hours, discharge during peak hours, and want to track the state of charge

I will use Hour only. Want to charge during 10-15 and discharge during 20-03 which is during the highest quantity used for A


```{r}
df_a <- df_a %>%
  mutate(
    action = case_when(
      Hour %in% 10:15 ~ "discharge",
      Hour %in% c(20:23, 0:3) ~ "charge",
      TRUE ~ "idle"
    )
  )
```

```{r}
df_a$charge_kwh    <- 0
df_a$discharge_kwh <- 0
df_a$soc           <- NA

soc <- 0.5 * battery_cap  # start at 50%

for (i in 1:nrow(df_a)) {
  act <- df_a$action[i]
  
  if (act == "charge" && soc < battery_cap) {
    charge <- min(1, battery_cap - soc)
    soc <- soc + charge
    df_a$charge_kwh[i] <- charge
  } else if (act == "discharge" && soc > 0) {
    discharge <- min(1, soc)
    soc <- soc - discharge
    df_a$discharge_kwh[i] <- discharge
  }
  
  df_a$soc[i] <- soc
}
```

Cost without battery:
```{r}
cost_no_batt_a <- sum(df_a$Quantity * df_a$Price, na.rm = TRUE)
```

```{r}
df_a <- df_a %>%
  mutate(
    grid_import = Quantity - discharge_kwh + charge_kwh,
    grid_import = pmax(grid_import, 0)
  )

cost_with_batt_a <- sum(df_a$grid_import * df_a$Price, na.rm = TRUE)

savings_a <- cost_no_batt_a - cost_with_batt_a
```

```{r}
print(cost_no_batt_a)
print(cost_with_batt_a)
```
```{r}
print(savings_a)
```

As we can see we are losing money with this model due to just going based on Quantity per hour instead of price.

Now want to do hours between 17 and 22 to discharge(5-10 PM) and between 0 and 5(12- 5 AM) which showed the highest price per hour for account A

```{r}
##Assign discharge and charge values
df_a <- df_a %>%
  mutate(
    action = case_when(
      Hour %in% 17:22 ~ "discharge",
      Hour %in% c(0:5) ~ "charge",
      TRUE ~ "idle"
    )
  )
```

For now I want to think about charging and discharging at 1kWh just to get the idea of what times would be good
```{r}
##SOC = State of Charge
df_a$charge_kwh    <- 0
df_a$discharge_kwh <- 0
df_a$soc           <- NA

soc <- 0.5 * battery_cap  # start at 50%

for (i in 1:nrow(df_a)) {
  act <- df_a$action[i]
  
  if (act == "charge" && soc < battery_cap) {
    charge <- min(1, battery_cap - soc)
    soc <- soc + charge
    df_a$charge_kwh[i] <- charge
  } else if (act == "discharge" && soc > 0) {
    discharge <- min(1, soc)
    soc <- soc - discharge
    df_a$discharge_kwh[i] <- discharge
  }
  
  df_a$soc[i] <- soc
}
```

Cost without battery:
```{r}
cost_no_batt_a <- sum(df_a$Quantity * df_a$Price, na.rm = TRUE)   ##Just quantity times price of the hour
```

Using Battery
```{r}
## Grid = Load-dishcarge + charge
df_a <- df_a %>%
  mutate(
    grid_import = Quantity - discharge_kwh + charge_kwh,
    grid_import = pmax(grid_import, 0)
  )

cost_with_batt_a <- sum(df_a$grid_import * df_a$Price, na.rm = TRUE)

savings_a <- cost_no_batt_a - cost_with_batt_a
```

```{r}
print(cost_no_batt_a)
print(cost_with_batt_a)
```
```{r}
print(savings_a)
```

When running on discharging with higher price, versus charging on low price we actually get some good savings.
